---
title: "CloudLUTpcaana"
author: "Michal Segal Rozenhaimer"
date: "Monday, July 06, 2015"
output: html_document
---

**This is a document that summarized the PCA analysis performed on the RSP simulated clouds LUT.
(R routine to perform this analysis:runpcaana)**

_read LUT and print some variable examples_

```{r,echo=FALSE}
        
        source('~/R/oracles_polar_algorithm/analysis_functions.R')
        print("processing NN_clouds_20150701.nc")

        # filepath: ~/R/oracles_polar_algorithm//data/NN_clouds_20150701.nc
        
        ## readncdf
        filename<-"~/R/oracles_polar_algorithm//data/NN_clouds_20150701.nc"
        ncin    <-readncdf(filename)

        #ncin <- open.ncdf("~/R/oracles_polar_algorithm//data/NN_clouds_20150701.nc")
        #print(ncin)
        filevarnames<-names(ncin$var)
        print(" file variable names:")
        print(filevarnames)

        ## get a sample of all vars in the nc file
        
        for (count in 1:length(filevarnames)) 
            {
              
              varname<-filevarnames[count]
              varval <- get.var.ncdf(ncin, filevarnames[count])
              nvar   <- dim(varval)
              print(paste("variable ", filevarnames[count], "dim:"))
              print(nvar)
              print(paste("variable ", filevarnames[count], "sample data:"))
              print(head(varval))
            }

        ## convert ncdf to data frame

        datframe<-rspLUT2dataframe(ncin)
        
        varval <- get.var.ncdf(ncin, "ref_i")
        dim2   <-  dim(varval)[2]
        dim3   <-  dim(varval)[3]
        
        close.ncdf(ncin)

```

_Show some example curtain plots of simulated cloud LUT_

##1. Total reflectance (I)

```{r, echo=FALSE}

## create melted subset

## create for ref_i
  
  #testname   <-names(datframe)[grepl("^ref_i_lambda", names(datframe))]
  #sub_refi_df<-melt(datframe,id.vars=c("case",filevarnames[1:6]),
  #                     measure.vars=testname,
  #                     variable_name="reflectance")

## plot heatmap using ggplot

  #ggplot(sub_refi_df, aes(x=reflectance,y=case, z= value)) + geom_tile(aes(fill= value)) + 
  #       scale_fill_gradient(low="blue", high="gold") + 
  #       theme(axis.text.x = element_blank(),
  #       axis.text.y = element_blank(),
  #       axis.ticks = element_blank()) + 
  #       ggtitle("total reflectance (I) for all 7 bands")

```

_subsetting reflectance (I) by lambda (n=7) and plotting curtains_

```{r, echo=FALSE}

## create melted subset

## create for ref_i

testname   <-names(datframe)[grepl("^ref_i_lambda", names(datframe))]
sub_refi_df<-melt(datframe,id.vars=c("case",filevarnames[1:6]),
                       measure.vars=testname,
                       variable_name="reflectance")

## add wavelength band factor

DT1 <- data.table(sub_refi_df)

## define factors

phrasein <-as.character()
phraseout<-as.character()

for (i in 1:dim2) {
  
  phrasein[i]  <- paste("^ref_i_lambda",as.character(i),sep="")
  phraseout[i] <- paste("lambda",as.character(i),sep="")
}

## assign factor lambda_i to each reflectance (ref_i) section

DT1[, lambda := ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[1], 
                                       names(datframe)))], phraseout[1],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[2], 
                                       names(datframe)))], phraseout[2],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[3], 
                                       names(datframe)))], phraseout[3], 
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[4], 
                                       names(datframe)))], phraseout[4],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[5], 
                                       names(datframe)))], phraseout[5],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[6], 
                                       names(datframe)))], phraseout[6], 
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[7], 
                                       names(datframe)))], phraseout[7], NA)))))))]

# all wavelength carpet plot for reflectance variables/split to 7 bands

ggplot(DT1, aes(x=reflectance,y=case, z= value)) + 
         geom_tile(aes(fill= value)) +
         facet_grid(.~ lambda, scales = "free", space = "free") +
         scale_fill_gradient(low="blue", high="gold", limits=c(0,1)) +
         # scale_fill_gradientn(colours=c("white","orange","red","dark red"), limits=c(-4,4))
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0.2, "lines")) +
         scale_x_discrete(expand = c(0, 0)) +
         scale_y_discrete(expand = c(0, 0)) +
         ggtitle("Total reflectance (I) splitted by 7 bands")

```

_subsetting reflectance (I) by lambda (n=7) and COD_

```{r, echo=FALSE}

# all wavelength carpet plot for reflectance variables/split to 7 bands

sp<-ggplot(DT1, aes(x=reflectance,y=as.factor(cod), z= value)) + 
         geom_tile(aes(fill= value)) +
         facet_grid(.~ lambda, scales = "free", space = "free") +
         #scale_fill_gradient(low="blue", high="gold", limits=c(0,1)) +
         scale_fill_gradientn(colours=c("light blue","blue",
                                          "light green","green",
                                          "pink","red","orange",
                                          "purple","gold"), limits=c(0,1)) +
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0, "lines"))
         # scale_x_discrete(expand = c(0, 0)) +
         # scale_y_discrete(expand = c(0, 0)) +
         sp + labs(title = "Total reflectance (I)", x = "VZA", y = "COD") +
         theme(axis.text.y = element_text(color = "black", size = 15, face = "italic"))
         #scale_y_discrete(limit = c("1", "2.5", "5","10","20")) +
         #scale_x_continuous(breaks = c(0,640*1,640*2,640*3,640*4,640*5,640*6,640*7),
         #                   labels = c("0","640","640","640","640","640","640","640"))
         

# change title, and x/y labels
# p.labs <- p + labs(title = "MAIN TITLE", x = "X-AXIS TITLE", y = "Y-AXIS TITLE")
# p.labs
# manipulate discrete axis labels
# scale_x_discrete(limit = c("I1", "SI2", "SI1"),labels = c("Ione","SItwo","SIone"))

```

_subsetting reflectance (I) by lambda (n=7) and Reff_

```{r, echo=FALSE}

# all wavelength carpet plot for reflectance variables/split to 7 bands

sp<-ggplot(DT1, aes(x=reflectance,y=as.factor(sizea), z= value)) + 
         geom_tile(aes(fill= value)) +
         facet_grid(.~ lambda, scales = "free", space = "free") +
         #scale_fill_gradient(low="blue", high="gold", limits=c(0,1)) +
         scale_fill_gradientn(colours=c("light blue","blue",
                                          "light green","green",
                                          "pink","red","orange",
                                          "purple","gold"), limits=c(0,1)) +
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0.2, "lines"))
         # scale_x_discrete(expand = c(0, 0)) +
         # scale_y_discrete(expand = c(0, 0)) +
         sp + labs(title = "Total reflectance (I)", x = "VZA", y = "Reff") +
         theme(axis.text.y = element_text(color = "black", size = 15, face = "italic"))

```

_subsetting reflectance (I) by lambda (n=7) and Azimuth angle_

```{r, echo=FALSE}

# all wavelength carpet plot for reflectance variables/split to 7 bands
windows()
sp<-ggplot(DT1, aes(x=reflectance,y=as.factor(azi), z= value)) + 
         geom_tile(aes(fill= value)) +
         facet_grid(.~ lambda, scales = "free", space = "free") +
         #scale_fill_gradient(low="blue", high="gold", limits=c(0,1)) +
         scale_fill_gradientn(colours=c("light blue","blue",
                                          "light green","green",
                                          "pink","red","orange",
                                          "purple","gold"), limits=c(0,1)) +
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0.2, "lines"))
         # scale_x_discrete(expand = c(0, 0)) +
         # scale_y_discrete(expand = c(0, 0)) +
         sp + labs(title = "Total reflectance (I)", x = "VZA", y = "AZ") +
         theme(axis.text.y = element_text(color = "black", size = 15, face = "italic"))

```

###1a. PCA on Total reflectance (I)

_PCA summary_
```{r,echo=FALSE}

## extract only total reflectance variables from datframe set

refivarnames <- names(datframe)[grepl("^ref_i_lambda", names(datframe))]

refi.dat     <- datframe[,refivarnames]
COD          <- as.factor(datframe$cod)
Reff         <- as.factor(datframe$sizea)
CTH          <- as.factor(datframe$cldtop)
SZA          <- as.factor(round(datframe$sza))
AZi          <- as.factor(datframe$azi)

pararray     <- data.frame(Reff = Reff, CTH = CTH, COD = COD, SZA = SZA, AZi = AZi)


## perform PCA with prcomp (on raw data)

refi.pca     <- prcomp(refi.dat)

## print PCA results

print("summary of PCA for total (I) reflectance:")

pcarefi.summary<-summary(refi.pca)

pcarefi.names  <-names(pcarefi.summary)

pcnumlogic     <- pcarefi.summary$importance[3,]<=0.99

print("number of PC's that represent 0.99 variance:")

pcarefi.summary$importance[,pcnumlogic==TRUE]

#print("pca variance plot")
#windows()
plot(refi.pca,type = "lines",
     main = "PC variance of total reflectance (I)")

print("most important PC's sdev:")

refi.pca$sdev[pcnumlogic==TRUE]

sampVZA   <- 140
sampVZAvec<- as.numeric()
for (i in 1:dim2){sampVZAvec[i]<-sampVZA + dim3*(i-1)}

print(paste("sample data rotation on PC space with VZA = ",
            as.character(sampVZA),sep=""))

refi.pca$rotation[sampVZAvec,pcnumlogic==TRUE]

## predict test

newset  <- rbind(head(refi.dat,2),tail(refi.dat,2))
predset <- predict(refi.pca,newset)

## perform princomp (with cov operation on data)
## need # cases > # of variables to perform
## refi.pca2 <- princomp(refi.dat, covmat = MASS::cov.rob(refi.dat))


```

_PCA plots_

```{r,echo=FALSE}

## try to visuallize PC's

refi.pca.df <- data.frame(refi.pca$rotation[,pcnumlogic==TRUE])
refi.pca.df <- cbind(refi.pca.df,vars=rownames(refi.pca.df))
pcnames     <-names(refi.pca.df)[grepl("^PC", names(refi.pca.df))]
sub_refi_df <-melt(refi.pca.df,id.vars=c("vars"),
                       measure.vars=pcnames,
                       variable_name="PC")

refi.pca.DT <- data.table(sub_refi_df)

## assign factor lambda_i to PC data table

refi.pca.DT[, lambda := ifelse(vars %in% names(datframe)[c(grepl(phrasein[1], 
                                       names(datframe)))], phraseout[1],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[2], 
                                       names(datframe)))], phraseout[2],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[3], 
                                       names(datframe)))], phraseout[3], 
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[4], 
                                       names(datframe)))], phraseout[4],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[5], 
                                       names(datframe)))], phraseout[5],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[6], 
                                       names(datframe)))], phraseout[6], 
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[7], 
                                       names(datframe)))], phraseout[7], NA)))))))]

## create PC curtain plots of lambda by VZA to inspect PC structure

sp<-ggplot(refi.pca.DT, aes(x=vars,y=PC, z= value)) + 
         geom_tile(aes(fill= value)) +
         scale_fill_gradientn(colours=c("blue","green",
                                        "pink","red","orange","gold")) +
         facet_grid(. ~ lambda, scales = "free", space = "free") +
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0.2, "lines")) +
         # scale_x_discrete(expand = c(0, 0)) +
         # scale_y_discrete(expand = c(0, 0)) +
         labs(title = "Total reflectance (I) PC loading map", x = "VZA", y = "") +
         theme(axis.text.y = element_text(color = "black", size = 15, face = "italic"))

# ggplot(df, aes(time,value)) + geom_line(aes(colour = series))
# or plot on different plots
# ggplot(df, aes(time,value)) + geom_line() + facet_grid(series ~ .)


## plot PC space with data
## windows()
## biplot(refi.pca)
```

_total reflectance by PC space_

```{r,echo=FALSE}
## create multiple biplots to explore the PC space

#  empty list for storage
gg_list <- list()

# select vars to plot for

varselect <- names(pararray)

# loop over varselect

  for(i in 1:length(varselect)) {
      
      # select specific variables to plot
      
      name   <- varselect[i]
      dattmp <- cbind(refi.dat,pararray[,i])
      names(dattmp)[dim(dattmp)[2]] <- varselect[i]
      
      # assign  ggplot to the i'th position in the list
      
      gg_list[[i]] <- autoplot(prcomp(refi.dat), data = dattmp, colour = name, label = TRUE)
      
      # print plot
      # print(gg_list[[i]])
      
  }
  
  ## print multiple variable
  #  multiplot(plotlist=gg_list,cols=3)
     multiplot(gg_list[[1]],gg_list[[3]],
               gg_list[[4]],gg_list[[5]],
               cols=2)







```

##2. Polarized Reflectance (Rp = sqrt(Q^2))

_subsetting Rp by lambda (n=7) and plotting curtains_

```{r, echo=FALSE}

## create melted subset

## create for Rp

testname   <-names(datframe)[grepl("^Rp_lambda", names(datframe))]
sub_Rp_df<-melt(datframe,id.vars=c("case",filevarnames[1:6]),
                       measure.vars=testname,
                       variable_name="reflectance")

## add wavelength band factor

DT2 <- data.table(sub_Rp_df)

## define factors

phrasein <-as.character()
phraseout<-as.character()

for (i in 1:dim2) {
  
  phrasein[i]  <- paste("^Rp_lambda",as.character(i),sep="")
  phraseout[i] <- paste("lambda",as.character(i),sep="")
}

## assign factor lambda_i to each Rp section

DT2[, lambda := ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[1], 
                                       names(datframe)))], phraseout[1],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[2], 
                                       names(datframe)))], phraseout[2],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[3], 
                                       names(datframe)))], phraseout[3], 
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[4], 
                                       names(datframe)))], phraseout[4],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[5], 
                                       names(datframe)))], phraseout[5],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[6], 
                                       names(datframe)))], phraseout[6], 
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[7], 
                                       names(datframe)))], phraseout[7], NA)))))))]

# all wavelength carpet plot for reflectance variables/split to 7 bands

ggplot(DT2, aes(x=reflectance,y=case, z= value)) + 
         geom_tile(aes(fill= value)) +
         facet_grid(.~ lambda, scales = "free", space = "free") +
         scale_fill_gradient(low="blue", high="gold", limits=c(0,0.1)) +
         # scale_fill_gradientn(colours=c("white","orange","red","dark red"), limits=c(-4,4))
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0.2, "lines")) +
         scale_x_discrete(expand = c(0, 0)) +
         scale_y_discrete(expand = c(0, 0)) +
         ggtitle("Polarized Reflectance (Rp) splitted by 7 bands")

```

###2a. PCA on Polarized reflectance (Rp)

_PCA summary_
```{r,echo=FALSE}

## extract only total reflectance variables from datframe set

refivarnames <- names(datframe)[grepl("^Rp_lambda", names(datframe))]

refq.dat     <- datframe[,refivarnames]
COD          <- as.factor(datframe$cod)
Reff         <- as.factor(datframe$sizea)
CTH          <- as.factor(datframe$cldtop)
SZA          <- as.factor(round(datframe$sza))
AZi          <- as.factor(datframe$azi)

pararray     <- data.frame(Reff = Reff, CTH = CTH, COD = COD, SZA = SZA, AZi = AZi)


## perform PCA with prcomp (on raw data)

refq.pca     <- prcomp(refq.dat)

## print PCA results

print("summary of PCA for Polarized reflectance (Rp):")

pcarefq.summary<-summary(refq.pca)

pcarefq.names  <-names(pcarefq.summary)

pcnumlogic     <- pcarefq.summary$importance[3,]<=0.99

print("number of PC's that represent 0.99 variance:")

pcarefq.summary$importance[,pcnumlogic==TRUE]

#print("pca variance plot")
#window()
plot(refq.pca,type = "lines",
     main = "PC variance of Polarized reflectance (Rp)")

print("most important PC's sdev:")

refq.pca$sdev[pcnumlogic==TRUE]

sampVZA   <- 140
sampVZAvec<- as.numeric()
for (i in 1:dim2){sampVZAvec[i]<-sampVZA + dim3*(i-1)}

print(paste("sample data rotation on PC space with VZA = ",
            as.character(sampVZA),sep=""))

refq.pca$rotation[sampVZAvec,pcnumlogic==TRUE]

## predict test

newset     <- rbind(head(refq.dat,2),tail(refq.dat,2))
predset_Rp <- predict(refq.pca,newset)

## perform princomp (with cov operation on data)
## need # cases > # of variables to perform
## refi.pca2 <- princomp(refi.dat, covmat = MASS::cov.rob(refi.dat))


```

_PCA plots_

```{r,echo=FALSE}

## try to visuallize PC's

refq.pca.df <- data.frame(refq.pca$rotation[,pcnumlogic==TRUE])
refq.pca.df <- cbind(refq.pca.df,vars=rownames(refq.pca.df))
pcnames     <-names(refq.pca.df)[grepl("^PC", names(refq.pca.df))]
sub_refq_df <-melt(refq.pca.df,id.vars=c("vars"),
                       measure.vars=pcnames,
                       variable_name="PC")

refq.pca.DT <- data.table(sub_refq_df)

## assign factor lambda_i to PC data table

refq.pca.DT[, lambda := ifelse(vars %in% names(datframe)[c(grepl(phrasein[1], 
                                       names(datframe)))], phraseout[1],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[2], 
                                       names(datframe)))], phraseout[2],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[3], 
                                       names(datframe)))], phraseout[3], 
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[4], 
                                       names(datframe)))], phraseout[4],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[5], 
                                       names(datframe)))], phraseout[5],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[6], 
                                       names(datframe)))], phraseout[6], 
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[7], 
                                       names(datframe)))], phraseout[7], NA)))))))]

## create PC curtain plots of lambda by VZA to inspect PC structure

sp<-ggplot(refq.pca.DT, aes(x=vars,y=PC, z= value)) + 
         geom_tile(aes(fill= value)) +
         scale_fill_gradientn(colours=c("blue","green",
                                        "pink","red","orange","gold")) +
         facet_grid(. ~ lambda, scales = "free", space = "free") +
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0.2, "lines")) +
         # scale_x_discrete(expand = c(0, 0)) +
         # scale_y_discrete(expand = c(0, 0)) +
         labs(title = "Polarized reflectance (Rp) PC loading map", x = "VZA", y = "") +
         theme(axis.text.y = element_text(color = "black", size = 15, face = "italic"))

# ggplot(df, aes(time,value)) + geom_line(aes(colour = series))
# or plot on different plots
# ggplot(df, aes(time,value)) + geom_line() + facet_grid(series ~ .)


## plot PC space with data
## windows()
## biplot(refi.pca)
```

_polarized reflectance by PC space_

```{r,echo=FALSE}
## create multiple biplots to explore the PC space

#  empty list for storage
gg_list <- list()

# select vars to plot for

varselect <- names(pararray)

# loop over varselect

  for(i in 1:length(varselect)) {
      
      # select specific variables to plot
      
      name   <- varselect[i]
      dattmp <- cbind(refq.dat,pararray[,i])
      names(dattmp)[dim(dattmp)[2]] <- varselect[i]
      
      # assign  ggplot to the i'th position in the list
      
      gg_list[[i]] <- autoplot(prcomp(refq.dat), data = dattmp, colour = name, label = TRUE)
      
      # print plot
      # print(gg_list[[i]])
      
  }
  
  ## print multiple variable
  #  multiplot(plotlist=gg_list,cols=3)
     multiplot(gg_list[[1]],gg_list[[3]],
               gg_list[[4]],gg_list[[5]],
               cols=2)







```

##3. Degree of Linear Polarization (DoLP = sqrt(Q^2))/I)

_subsetting DoLP by lambda (n=7) and plotting curtains_

```{r, echo=FALSE}

## create melted subset

## create for DoLP

testname   <-names(datframe)[grepl("^DoLP_lambda", names(datframe))]
sub_DoLP_df<-melt(datframe,id.vars=c("case",filevarnames[1:6]),
                       measure.vars=testname,
                       variable_name="reflectance")

## add wavelength band factor

DT3 <- data.table(sub_DoLP_df)

## define factors

phrasein <-as.character()
phraseout<-as.character()

for (i in 1:dim2) {
  
  phrasein[i]  <- paste("^DoLP_lambda",as.character(i),sep="")
  phraseout[i] <- paste("lambda",as.character(i),sep="")
}

## assign factor lambda_i to each DoLP section

DT3[, lambda := ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[1], 
                                       names(datframe)))], phraseout[1],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[2], 
                                       names(datframe)))], phraseout[2],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[3], 
                                       names(datframe)))], phraseout[3], 
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[4], 
                                       names(datframe)))], phraseout[4],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[5], 
                                       names(datframe)))], phraseout[5],
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[6], 
                                       names(datframe)))], phraseout[6], 
               ifelse(reflectance %in% names(datframe)[c(grepl(phrasein[7], 
                                       names(datframe)))], phraseout[7], NA)))))))]

# all wavelength carpet plot for reflectance variables/split to 7 bands

ggplot(DT3, aes(x=reflectance,y=case, z= value)) + 
         geom_tile(aes(fill= value)) +
         facet_grid(.~ lambda, scales = "free", space = "free") +
         scale_fill_gradient(low="blue", high="gold", limits=c(0,0.5)) +
         # scale_fill_gradientn(colours=c("white","orange","red","dark red"), limits=c(-4,4))
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0.2, "lines")) +
         scale_x_discrete(expand = c(0, 0)) +
         scale_y_discrete(expand = c(0, 0)) +
         ggtitle("DoLP splitted by 7 bands")

```
_subsetting DoLP by lambda (n=7) and COD_

```{r, echo=FALSE}

# all wavelength carpet plot for reflectance variables/split to 7 bands

sp<-ggplot(DT3, aes(x=reflectance,y=as.factor(cod), z= value)) + 
         geom_tile(aes(fill= value)) +
         facet_grid(.~ lambda, scales = "free", space = "free") +
         #scale_fill_gradient(low="blue", high="gold", limits=c(0,1)) +
         scale_fill_gradientn(colours=c("light blue","blue",
                                          "light green","green",
                                          "pink","red","orange",
                                          "purple","gold"), limits=c(0,0.5)) +
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0.2, "lines"))
         # scale_x_discrete(expand = c(0, 0)) +
         # scale_y_discrete(expand = c(0, 0)) +
         sp + labs(title = "DoLP", x = "VZA", y = "COD") +
         theme(axis.text.y = element_text(color = "black", size = 15, face = "italic"))

```

_subsetting DoLP by lambda (n=7) and Reff_

```{r, echo=FALSE}

# all wavelength carpet plot for reflectance variables/split to 7 bands

sp<-ggplot(DT3, aes(x=reflectance,y=as.factor(sizea), z= value)) + 
         geom_tile(aes(fill= value)) +
         facet_grid(.~ lambda, scales = "free", space = "free") +
         #scale_fill_gradient(low="blue", high="gold", limits=c(0,1)) +
         scale_fill_gradientn(colours=c("light blue","blue",
                                          "light green","green",
                                          "pink","red","orange",
                                          "purple","gold"), limits=c(0,0.5)) +
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0.2, "lines"))
         # scale_x_discrete(expand = c(0, 0)) +
         # scale_y_discrete(expand = c(0, 0)) +
         sp + labs(title = "DoLP", x = "VZA", y = "Reff") +
         theme(axis.text.y = element_text(color = "black", size = 15, face = "italic"))

```

_subsetting DoLP by lambda (n=7) and Azimuth angle_

```{r, echo=FALSE}

# all wavelength carpet plot for reflectance variables/split to 7 bands

sp<-ggplot(DT3, aes(x=reflectance,y=as.factor(azi), z= value)) + 
         geom_tile(aes(fill= value)) +
         facet_grid(.~ lambda, scales = "free", space = "free") +
         #scale_fill_gradient(low="blue", high="gold", limits=c(0,1)) +
         scale_fill_gradientn(colours=c("light blue","blue",
                                          "light green","green",
                                          "pink","red","orange",
                                          "purple","gold"), limits=c(0,0.5)) +
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0.2, "lines"))
         # scale_x_discrete(expand = c(0, 0)) +
         # scale_y_discrete(expand = c(0, 0)) +
         sp + labs(title = "DoLP", x = "VZA", y = "AZ") +
         theme(axis.text.y = element_text(color = "black", size = 15, face = "italic"))

```

###3a. PCA on Degree of linear polarization (DoLP)

_PCA summary_
```{r,echo=FALSE}

## extract only total reflectance variables from datframe set

refivarnames <- names(datframe)[grepl("^DoLP_lambda", names(datframe))]

refd.dat     <- datframe[,refivarnames]
COD          <- as.factor(datframe$cod)
Reff         <- as.factor(datframe$sizea)
CTH          <- as.factor(datframe$cldtop)
SZA          <- as.factor(round(datframe$sza))
AZi          <- as.factor(datframe$azi)

pararray     <- data.frame(Reff = Reff, CTH = CTH, COD = COD, SZA = SZA, AZi = AZi)


## perform PCA with prcomp (on raw data)

refd.pca     <- prcomp(refd.dat)

## print PCA results

print("summary of PCA for DoLP:")

pcarefd.summary<-summary(refd.pca)

pcarefd.names  <-names(pcarefd.summary)

pcnumlogic     <- pcarefd.summary$importance[3,]<=0.99

print("number of PC's that represent 0.99 variance:")

pcarefd.summary$importance[,pcnumlogic==TRUE]

#print("pca variance plot")
#window()
plot(refd.pca,type = "lines",
     main = "PC variance of DoLP")

print("most important PC's sdev:")

refd.pca$sdev[pcnumlogic==TRUE]

sampVZA   <- 140
sampVZAvec<- as.numeric()
for (i in 1:dim2){sampVZAvec[i]<-sampVZA + dim3*(i-1)}

print(paste("sample data rotation on PC space with VZA = ",
            as.character(sampVZA),sep=""))

refd.pca$rotation[sampVZAvec,pcnumlogic==TRUE]

## predict test

newset       <- rbind(head(refd.dat,2),tail(refd.dat,2))
predset_DoLP <- predict(refd.pca,newset)

## perform princomp (with cov operation on data)
## need # cases > # of variables to perform
## refi.pca2 <- princomp(refi.dat, covmat = MASS::cov.rob(refi.dat))


```

_PCA plots_

```{r,echo=FALSE}

## try to visuallize PC's

refd.pca.df <- data.frame(refd.pca$rotation[,pcnumlogic==TRUE])
refd.pca.df <- cbind(refd.pca.df,vars=rownames(refd.pca.df))
pcnames     <-names(refd.pca.df)[grepl("^PC", names(refd.pca.df))]
sub_refd_df <-melt(refd.pca.df,id.vars=c("vars"),
                       measure.vars=pcnames,
                       variable_name="PC")

refd.pca.DT <- data.table(sub_refd_df)

## assign factor lambda_i to PC data table

refd.pca.DT[, lambda := ifelse(vars %in% names(datframe)[c(grepl(phrasein[1], 
                                       names(datframe)))], phraseout[1],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[2], 
                                       names(datframe)))], phraseout[2],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[3], 
                                       names(datframe)))], phraseout[3], 
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[4], 
                                       names(datframe)))], phraseout[4],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[5], 
                                       names(datframe)))], phraseout[5],
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[6], 
                                       names(datframe)))], phraseout[6], 
                ifelse(vars %in% names(datframe)[c(grepl(phrasein[7], 
                                       names(datframe)))], phraseout[7], NA)))))))]

## create PC curtain plots of lambda by VZA to inspect PC structure

sp<-ggplot(refd.pca.DT, aes(x=vars,y=PC, z= value)) + 
         geom_tile(aes(fill= value)) +
         scale_fill_gradientn(colours=c("blue","green",
                                        "pink","red","orange","gold")) +
         facet_grid(. ~ lambda, scales = "free", space = "free") +
         theme(axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         panel.margin = unit(0.2, "lines")) +
         # scale_x_discrete(expand = c(0, 0)) +
         # scale_y_discrete(expand = c(0, 0)) +
         labs(title = "DoLP PC loading map", x = "VZA", y = "") +
         theme(axis.text.y = element_text(color = "black", size = 15, face = "italic"))

# ggplot(df, aes(time,value)) + geom_line(aes(colour = series))
# or plot on different plots
# ggplot(df, aes(time,value)) + geom_line() + facet_grid(series ~ .)


## plot PC space with data
## windows()
## biplot(refi.pca)
```

_DoLP by PC space_

```{r,echo=FALSE}
## create multiple biplots to explore the PC space

#  empty list for storage
gg_list <- list()

# select vars to plot for

varselect <- names(pararray)

# loop over varselect

  for(i in 1:length(varselect)) {
      
      # select specific variables to plot
      
      name   <- varselect[i]
      dattmp <- cbind(refd.dat,pararray[,i])
      names(dattmp)[dim(dattmp)[2]] <- varselect[i]
      
      # assign  ggplot to the i'th position in the list
      
      gg_list[[i]] <- autoplot(prcomp(refd.dat), data = dattmp, colour = name, label = TRUE)
      
      # print plot
      # print(gg_list[[i]])
      
  }
  
  ## print multiple variable
  #  multiplot(plotlist=gg_list,cols=3)
     multiplot(gg_list[[1]],gg_list[[3]],
               gg_list[[4]],gg_list[[5]],
               cols=2)







```




